### 2015-05-09 papers/histogram/thesis/thesis.tex
made a fix in non-vanilla wang landau for fair comparison

We find that Wang-Landau is capable of performing the best of all
methods, but fails catastrophically without a predefined energy range
for the square-well fluid. Our first method gives results comparable
to successful Wang-Landau simulations, but with some anomalous
cases. Our implementation of the optimized ensemble using a transition
matrix results in worse performance over the straightforward
transition matrix Monte Carlo method.

@@

Once we identify an energy $E_-$ satisfying (\ref{eq:min_E_condition})
in an initialization routine for any histogram method, we effectively
forbid transitions to energies $E<E_-$ by tentatively setting weights
$w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ for some temperature $T\ll
T_{\t{min}}$. These weights prevent the initialization routine from
wasting time on energies at which we will override the weight array
after initialization anyways. We always set
$w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
histogram methods with $kT_{\t{min}}/\epsilon=0.2$. We chose this
temperature because it is right below the temperature at which the
square-well fluid condenses from a gas into a liquid (see Section
\ref{sec:properties}). \fixme{what about pressure?}

We made a special case for Wang-Landau, as the design of this method
assumes knowledge of the energy range of interest a priori. In order
to more fairly compare Wang-Landau to other methods, we therefore
simulated two cases: one in which the minimum important energy $E_-$
is identified via other methods and hard-coded into the Wang-Landau
initialization routine, and one in which Wang-Landau simply assumes
that the lowest energy a given simulation has ever observed is the
minimum important energy. The simulations in which $E_-$ is hard-coded
we will refer to as ``vanilla Wang-Landau'' simulations.

>>>>

We find that Wang-Landau potentially performs better than other
histogram methods, but fails catastrophically without a predefined
energy range for the square-well fluid. Our first method gives results
comparable to successful Wang-Landau simulations, but can have
anomalous cases with large error. The transition matrix Monte Carlo
consistently yields results whose quality is marginally worse than
functional Wang-Landau. Finally, our implementation of the optimized
ensemble using a transition matrix results in worse performance over
the straightforward transition matrix Monte Carlo method.

@@

Once we identify an energy $E_-$ satisfying (\ref{eq:min_E_condition})
in an initialization routine for any histogram method, we set weights
$w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ with a temperature
$kT_{\t{min}}/\epsilon=0.2$. These weights prevent the initialization
routine from wasting time on energies at which we will override the
weight array after initialization anyways. We chose this temperature
because it is comfortably below the temperature at which the
square-well fluid condenses from a gas into a liquid (see Section
\ref{sec:properties}). \fixme{what about pressure?}

We made a special case for Wang-Landau treatment of the energy range
of interest in, as the design of this method assumes a priori
knowledge of $E_-$. In order to more fairly compare Wang-Landau to
other methods, we therefore also determined the appropriate energy
$E_-$ for each system size $N$ using other histogram methods, and
simulated via Wang-Landau with the a lower energy bound hard-coded. We
will distinguish the two sets of Wang-Landau simulations by referring
to those in which $E_-$ is hard-coded as ``vanilla Wang-Landau.''
